Convex Surrogates via Polyhedral Losses

### 2019-02-01 cvx-fin.tex
Make notation consistent

Let $v_1,\dots,v_5$ be any $5$ different vectors in $\reals^2$, rows of a matrix $V$.
Let $S_y = \{w : Vw \leq B_y\}$ for each column $B_y$ of $B$, $y\in\{1,\dots,5\}$.
Recall we wish to prove that not all H-reps $(V,B_y)$ are tight, or in other words, there is some strictly loose constraint in one of the definitions of $S_y$.

WLOG, suppose $v_1 = (0,1)$ and they are numbered clockwise with $\|v_1\| \geq \|v_j\| (\forall j \neq i)$.
Let $\theta_{i,j}$ be the clockwise angle between $v_i$ and $v_j$.
The vectors point to the sides of an irregular pentagon $S^* = \{w : Vw \leq \vec{1}\}$.
Let $w_{1,2}$ be the corner of this pentagon between $v_1$ and $v_2$, and so on.
In our case, each $S_y$ is the intersection of $S^*$ with a halfspace, $S_y = \{w \in S^* : \inprod{v_y}{w} \leq -1\}$.
So in particular, $(V,B_1)$ is tight only if all three vertices $w_{23},w_{34},w_{45}$ satisfy $\inprod{v_1}{w} \leq -1$.
(If for example $\inprod{v_1}{w_{23}} > -1$, then the constraint $\inprod{v_2}{w} \leq 1$ is not tight, i.e. there is a separation between it and $S_1$.)
Because $v_1$ is the largest vector by assumption, this implies $\theta_{13} \geq 180^{\circ}$, but it also implies $\theta_{41} \geq 180^{\circ}$, which is a contradiction (both vectors would have to lie directly opposite $v_1$, but then they would be equal).
\end{proof}

@@

\paragraph{On the matching upper bound.}
In constrast to this lower bound, \bo{cite} give a $2$-embedding of the abstain loss on $|\Y|=4$.
This implies that a $2$-construction for the $4 \times 4$ version of the above $B$ matrix exists.
Indeed, one can take $V$ to be the four unit vectors pointing directly north, south, east, and west; the H-representations are then line segments, e.g. the segment $w_1=-1, w_2 \in [-1,1]$, etc.
These are all tight.
Indeed, as required by our characterization, the subgradient sets of their loss functions at the abstain embedding point span these H-reps (in fact are equal to them).

@@

\begin{theorem}
  Abstain with $\alpha > 1/2$ and $|\Y|=7$ is not $2$-embeddable.
\end{theorem}
\begin{proof}
  Like Theorem~\ref{thm:abs-5outcome-not-2embed}, we will again look at the abstain cell.  
  First, note that when $\alpha \geq \frac{n-1}{n}$, the abstain property decomposes into the mode, as the abstain cell is empty.
  Given $\frac{n-1}{n} > \alpha > 1/2$, we can write the matrix $B$ describing the abstain cell as follows:
	\[ B = \begin{pmatrix}
       -\beta &  1 &  1 & \cdots & 1  \\
       1  & -\beta &  1 & \cdots & 1  \\
          &    &    & \vdots &    \\
       1  &  1 &  1 & \cdots & -\beta
     \end{pmatrix} .
  \]

  We know that $\beta \in (0,1)$ as the inequalities are strict.
\end{proof}

>>>>

Let $v_1,\dots,v_5$ be any $5$ different vectors in $\reals^2$, rows of a matrix $V$.
Let $T_y = \{w : Vw \leq B_y\}$ for each column $B_y$ of $B$, $y\in\{1,\dots,5\}$.
Recall we wish to prove that not all H-reps $(V,B_y)$ are tight, or in other words, there is some strictly loose constraint in one of the definitions of $T_y$.

WLOG, suppose $v_1 = (0,1)$ and they are numbered clockwise with $\|v_1\| \geq \|v_j\| (\forall j \neq i)$.
Let $\theta_{i,j}$ be the clockwise angle between $v_i$ and $v_j$.
The vectors point to the sides of an irregular pentagon $T^* = \{w : Vw \leq \vec{1}\}$.
Let $w_{1,2}$ be the corner of this pentagon between $v_1$ and $v_2$, and so on.
In our case, each $S_y$ is the intersection of $T^*$ with a halfspace, $T_y = \{w \in T^* : \inprod{v_y}{w} \leq -1\}$.
So in particular, $(V,B_1)$ is tight only if all three vertices $w_{23},w_{34},w_{45}$ satisfy $\inprod{v_1}{w} \leq -1$.
(If for example $\inprod{v_1}{w_{23}} > -1$, then the constraint $\inprod{v_2}{w} \leq 1$ is not tight, i.e. there is a separation between it and $T_1$.)
Because $v_1$ is the largest vector by assumption, this implies $\theta_{13} \geq 180^{\circ}$, but it also implies $\theta_{41} \geq 180^{\circ}$, which is a contradiction (both vectors would have to lie directly opposite $v_1$, but then they would be equal).
\end{proof}

@@

\paragraph{On the matching upper bound.}
In constrast to this lower bound, \cite{ramaswamy2018consistent} give a $2$-embedding of the abstain loss on $|\Y|=4$.
This implies that a $2$-construction for the $4 \times 4$ version of the above $B$ matrix exists.
Indeed, one can take $V$ to be the four unit vectors pointing directly north, south, east, and west; the H-representations are then line segments, e.g. the segment $w_1=-1, w_2 \in [-1,1]$, etc.
These are all tight.
Indeed, as required by our characterization, the subgradient sets of their loss functions at the abstain embedding point span these H-reps (in fact are equal to them).

@@

\begin{theorem}
  Abstain with $\alpha > 1/2$ and $|\Y|=7$ is not $2$-embeddable.
\end{theorem}
\begin{proof}
  Like Theorem~\ref{thm:abs-5outcome-not-2embed}, we will again look at the abstain cell.  
  First, note that when $\alpha \geq \frac{n-1}{n}$, the abstain property decomposes into the mode, as the abstain cell is empty, so we can restrict to $\frac{n-1}{n} > \alpha > 1/2$.
  We then write the matrix $B$ describing the abstain cell as follows:
	\[ B = \begin{pmatrix}
       -\beta &  1 &  1 & \cdots & 1  \\
       1  & -\beta &  1 & \cdots & 1  \\
          &    &    & \vdots &    \\
       1  &  1 &  1 & \cdots & -\beta
     \end{pmatrix} .
  \]

  We know that $\beta := 1 - \frac{1}{\alpha} \in (0,1)$ as the inequalities are strict.
  Consider two requirements of the polytope: first, for all $i \in \{1,2, \ldots, 6\}, ||v_i|| > 0$.
  Additionally, we know $-\beta < 0$.
  Therefore, there will always be some $S_i$ such that $0 \not \in S_i$.

  By a similar argument to Theorem~\ref{thm:abs-5outcome-not-2embed}, we then must observe $\theta_{j, j+3} > 180^o$ in order to have some $w$ that satisfies our tightness constraint for $j+3$.
  That is, there is an $x \in S_i$ such that $\inprod{v_{j+3}}{x} = 1$.
  However, this implies $\theta_{j+3, j} < 180^o$, and therefore, there is no $x$ such that $\inprod{v_j}{x} = 1$. 
\end{proof}
